---
title: "PRJNA767017 processing"
output: 
  html_document:
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, cache.lazy=FALSE, message=FALSE, warning=FALSE)
library(sctransform)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(cowplot)
library(ggrepel)
library(dplyr)
library(tidyr)
options(future.globals.maxSize = 10 * 1024 * 1024^2)
```

we follow the [recent paper](https://zenodo.org/record/7555405) and use their data as reference:

```{r get_reference, fig.width=6,fig.height=5}
reference <- readRDS(file.path('seurat',"PBMC_vaccine_CITE.rds"))
```

load PBMC data from [PRJNA767017](https://insight.jci.org/articles/view/153201); take all cells with less than 10% mitochondrial RNA and between 250 and 5000 genes

- Vacc_pre (4 donors)
- Vacc_post (4 donors)
- HCW_base (3 donors)
- HCW_conv (3 donors)

```{r get_data}
samples <- c("Vacc_pre","Vacc_post","HCW_base","HCW_conv")
ndonors <- c(4,4,3,3)
names(ndonors) <- samples

pbmc <- list()
for (sample in samples) {
  dge <- Read10X_h5(file.path('cellranger',paste0(sample,'filtered_counts.h5'))
  sobj <- CreateSeuratObject(counts = dge[['Gene Expression']], project=sample)
  sobj[['pct.mito']] <- PercentageFeatureSet(sobj, pattern='^MT-')
  sobj <- subset(sobj, (pct.mito < 10) & (nFeature_RNA <= 5000) & (nFeature_RNA >= 250))
  capture.output(sobj <- SCTransform(sobj, verbose = FALSE),file='/dev/null')
  cells <- Cells(sobj)
  hts <- row.names(dge[['Antibody Capture']])[order(rowSums(dge[['Antibody Capture']]),decreasing=TRUE)[1:ndonors[[sample]]]]
  sobj[["HTO"]] <- CreateAssayObject(counts = dge[['Antibody Capture']][hts,cells,drop=FALSE])
  sobj <- NormalizeData(sobj, normalization.method='CLR', assay='HTO', verbose=FALSE)
  sobj <- HTODemux(sobj, assay='HTO', verbose=FALSE)
  vireo <- read.table(file.path('vireo',sample,'donor_ids.tsv'),
                      header=1,row.names=1,as.is=TRUE)
  colnames(vireo) <- paste0('vireo_',colnames(vireo))
  sobj <- AddMetaData(sobj, vireo) 
  print(sample)
  print(table(sobj$hash.ID,sobj$vireo_donor_id))
  cells <- Cells(sobj)[grepl('TotalSeq',sobj$hash.ID) & grepl('donor',sobj$vireo_donor_id)]
  sobj <- subset(sobj, cells=cells)
  pbmc[[sample]] <- sobj
}
```

get some QC metrics

```{r QC,fig.width=10,fig.height=5}
md <- do.call(rbind, lapply(names(pbmc), function(x) pbmc[[x]]@meta.data[,c('nCount_RNA','nFeature_RNA','pct.mito')] %>%
                              dplyr::mutate(orig.ident=x))) %>%
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(orig.ident=paste0(orig.ident,' (n=',n(),')')) %>%
  gather(metric,value,-orig.ident)

ggplot(md,aes(x=orig.ident,y=value,fill=orig.ident)) + 
  geom_violin(scale='width') +
  facet_wrap(~metric,scales='free_y') + 
  theme_classic() + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none') + 
  labs(x='',y='')
```

project our cells onto this reference embedding using RNA data; predict cell type labels for the level 1, level 2 and level 3 clustering of the reference

```{r project}
for (sample in samples) {
  anchors <- FindTransferAnchors(
    reference = reference,
    query = pbmc[[sample]],
    normalization.method = "SCT",
    reference.reduction = "spca",
    dims = 1:50,
    k.anchor = 5,
    verbose=FALSE)
  pbmc[[sample]] <- MapQuery(
    anchorset = anchors,
    query = pbmc[[sample]],
    reference = reference,
    refdata = list(
      celltype.l1 = "celltypel1",
      celltype.l2 = "celltypel2",
      celltype.l3 = "celltypel3"
    ),
    reference.reduction = "spca", 
    reduction.model = "wnn.umap",
    verbose=FALSE)
}
```

identify and remove putative doublets (shown in pink), and cells with prediction score < .75

```{r doublets,fig.width=10,fig.height=12}
source(file.path('R','my_doubletFinder.R'))
pbmc <- lapply(pbmc, function(sobj) {
  DefaultAssay(sobj) <- 'RNA'
  Idents(sobj) <- 'predicted.celltype.l1'
  homotypic.prop <- modelHomotypic(Idents(sobj))
  nExp_poi <- round(0.05*nrow(sobj@meta.data)) 
  nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
  tmp <- my_doubletFinder(sobj, PCs = 1:10, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE)
  tmp <- my_doubletFinder(tmp, PCs = 1:10, pN = 0.25, pK = 0.09, nExp = nExp_poi.adj, reuse.pANN = "pANN")
  AddMetaData(sobj, tmp@meta.data[,'DF.classifications',drop=FALSE])
})

do.call(rbind,lapply(pbmc, function(sobj) table(sobj$DF.classifications))) %>%
  DT::datatable()

plots <- lapply(names(pbmc), function(sample)
  DimPlot(pbmc[[sample]], group.by='DF.classifications',
          cols=c('Singlet'='gray',
                 'Doublet'='deeppink')) +
    theme_void() + 
    labs(title=sample) + 
    theme(legend.position='none'))

plot_grid(plotlist=plots,ncol=5,align='vh')

pbmc.sub <- lapply(pbmc, function(x) subset(x, subset=(DF.classifications=='Singlet') & (predicted.celltype.l1.score >= .75)))
```

```{r save}
for (sample in samples) {
  saveRDS(pbmc.sub[[sample]], file.path('seurat',paste0(sample,'.rds')))
}
```

```{r sessionInfo}
sessionInfo()
```

